---
title: "ZebraFinches"
author: "Theresa Michalke"
date: "6/11/2021"
output:
  html_document:
    code_folding: hide
    css: bootstrap.min.css
    df_print: paged
    number_sections: yes
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
library(tidyverse)
library(stringr)
library(ggplot2)
library(dplyr)
library(viridis)
library(hrbrthemes)
library(plotly)
library(ggbeeswarm)

knitr::opts_chunk$set(echo = TRUE)

```
# Introduction
Our data is a collection of agonistic interactions and locations of individual finches in a grid of 5x5. 
Measurements were collected in multiple sessions. Each pair of finches nested in the following locations.
```{r nest, echo=FALSE}
nest<-matrix(0,ncol=3,nrow=25)
nest<-as.data.frame(nest)
colnames(nest)<-c("x","y","nest")
i<-1
for(x in c(1:5)){
  for (y in c(1:5)){
    nest[i,1]<-x
    nest[i,2]<-y
    nest[i,3]<-"empty"
    i<-i+1
  }
}
nest[10,3]<-"lamp"
nest[4,3]<-"nest_b"
nest[2,3]<-"nest_c"
ggplot(nest,aes(x=x,y=y,fill=nest))+geom_tile( colour = "grey50")+scale_fill_viridis(discrete=TRUE)+theme_minimal() +theme(legend.position = 'top')+
  guides(fill = guide_legend(title.position = 'top', title.hjust = .5, barwidth = unit(20, 'lines'), barheight = unit(.5, 'lines')))
```

Our analysis consists first on creating and cleaning the data, followed by some visualizations and statistical tests. 

## Reading Data
Our data is in a csv format, with the following columns
```{r summary}
my_pile_of_stuff<-read.csv("Birdies_exp.csv") #("Zebra_finch_exp.csv")
my_pile_of_stuff<-as.data.frame(my_pile_of_stuff)
#my_pile_of_stuff
```
This data needs to be formatted in the following ways:

*  Data needs to be separated between interactions and movement measurements
*  The variable territory denotes the location in a grid with a letter and a number. This needs to be split up and put in an X-Y numeric coordinate system.
*  The location of each birds nest is known, but wasnt added to the data. This is important because in the agonistic interactions, the _treatment_ is where the food was places, and this corresponds to a nest location
*  During the agonistic interactions, we consider a treatment where the food was place, which overlapped with a pair of birds nest location, due to where the birds decided to nest, 2 nests were in close proximity, meaning that depending on the condition we relabelled treatment by proximity to the nest.
*  We also want to calculate the distance between an interaction and the proximity to the food


```{r read}
#We obtain the names of the birds, and then map these values to their corresponding nest
birdies<-levels(factor(my_pile_of_stuff$identity))
nests<-c("nest_b","nest_b","lamp","nest_c","nest_c","lamp")
my_pile_of_stuff$nest<-factor(plyr::mapvalues(my_pile_of_stuff$identity,birdies,nests))
my_pile_of_stuff$exp_id<-paste0(my_pile_of_stuff$day,"-",my_pile_of_stuff$session)

#Depending on the treatment, the food location was either close or far from each nest, so we create a new variable to show this
my_pile_of_stuff['closeness']<-"far"
my_pile_of_stuff[my_pile_of_stuff$treatment == my_pile_of_stuff$nest,"closeness"]<-"close"
my_pile_of_stuff[my_pile_of_stuff$treatment=="lamp" & my_pile_of_stuff$nest=="nest_b","closeness"]<-"close"
my_pile_of_stuff[my_pile_of_stuff$treatment =="nest_b" & my_pile_of_stuff$nest=="lamp","closeness"]<-"close"

```
This part takes care of the movement table, as well as making plots with the variable of interest
```{r movement_data,message=FALSE,warning=FALSE}
#We select only rows where territory is measured (this is the measurements of where the finch was in a given time)
movement<-my_pile_of_stuff[!is.na(my_pile_of_stuff$territory),]
movement<-movement[,c(-10,-11)]
tmp<-str_split_fixed(movement$territory, "",n=2)
movement$x_pos<-tmp[,1]
#The way that data was captured, some fields are empty and we convert them to NA
movement$x_pos[movement$x_pos==""]<-NA
#converts the number part into a factor
movement$x_pos<-as.factor(movement$x_pos)
movement$x_pos<-droplevels(movement$x_pos)
movement$x_pos<-as.numeric(movement$x_pos)
#makes the leters into a factor, and converts them to a numeric variable
movement$y_pos<-as.factor(tmp[,2])
movement$y_pos[movement$y_pos==""]<-NA
movement$y_pos<-as.factor(movement$y_pos)
movement$y_pos<-droplevels(movement$y_pos)
movement$y_pos<-as.numeric(movement$y_pos)
#Our setup started numbering the grid from top to bottom, we transform the numbers so visualizations match the experimental set up
#Only taking into account the rows with a territory
movement$y_pos<-movement$y_pos *-1+6
range<-movement[complete.cases(movement$territory),]

#We create a function that takes as an input a variable, and creates a heatmap of the counts of the position of each bird across the grid #normalized by the total amount of recorded events
plot_heat<-function(df,cond){
  #print(cond)
  hm<-df %>% dplyr::count(identity, x_pos,y_pos,df[cond], sort = TRUE)
  #print(hm)
  nd<-df %>% dplyr::count(identity,df[cond], sort = TRUE)
  for(x in 1:dim(nd)[1]){
    hm[hm$identity==nd[x,"identity"] & hm[cond]==nd[x,cond],"n"] <- hm[hm$identity==nd[x,"identity"] & hm[cond]==nd[x,cond],"n"]/nd[x,"n"]
  }
  p<-ggplot(hm,aes(x=x_pos,y=y_pos,fill=n))+geom_tile()+facet_grid(as.formula(paste(cond,"~", "identity")))+scale_fill_viridis(discrete=FALSE) + theme_bw()+theme(plot.title.position = 'plot')+theme(legend.position = 'top')+guides(fill = guide_colorbar(title="Position counts",title.position = 'top', title.hjust = .5,
 barwidth = unit(20, 'lines'), barheight = unit(.5, 'lines'))) +
    labs(title = paste0("Movement per ",cond), xlab= "x position",ylab= "y position")
  print(p)
  return(hm)
}
```

We first visualize the positions of the finches depending on the day

```{r movement_plot,message=FALSE,warning=FALSE,fig.width=7, fig.height=14}
tmp3<-plot_heat(range,"day") #enter whatever variable
```

We also look at the positions dependent on other variables

*  cycle
*  weather
```{r movement_plot2,message=FALSE,warning=FALSE}
tmp4<-plot_heat(range,"cycle") 
tmp5<-plot_heat(range,"weather") 
#dev.off()
#for(x in 1:dim(tmp2)[1]){
#  print(x)
#}
```
This part takes care of the interaction table, as well as calculating the distances
```{r interaction,,message=FALSE,warning=FALSE}
interaction<-my_pile_of_stuff[,-9]
interaction$treatment<-as.factor(interaction$treatment)

interaction<-interaction[complete.cases(interaction$agonistic),]

tmp<-str_split_fixed(interaction$agonistic, "",n=2)
interaction$x_pos<-tmp[,1]
#The way that data was captured, some fields are empty and we convert them to NA
interaction$x_pos[interaction$x_pos==""]<-NA

#converts the number part into a factor
interaction$x_pos<-as.factor(interaction$x_pos)
interaction$x_pos<-droplevels(interaction$x_pos)
interaction$x_pos<-as.numeric(interaction$x_pos)

#makes the leters into a factor, and converts them to a numeric variable
interaction$y_pos<-as.factor(tmp[,2])
interaction$y_pos[interaction$y_pos==""]<-NA
interaction$y_pos<-as.factor(interaction$y_pos)
interaction$y_pos<-droplevels(interaction$y_pos)
interaction$y_pos<-as.numeric(interaction$y_pos)

#Our setup started numbering the grid from top to bottom, we transform the numbers so visualizations match the experimental set up
interaction$y_pos<-interaction$y_pos *-1+6

#we create a data frame with the coordinates of the food (and therefore also the nest)
nest_coord<-data.frame(matrix(nrow = 2,ncol=4))
colnames(nest_coord)<-levels(interaction$treatment)
nest_coord[1,1]<-3
nest_coord[2,1]<-3
nest_coord[1,2]<-2
nest_coord[2,2]<-5
nest_coord[1,3]<-1
nest_coord[2,3]<-4
nest_coord[1,4]<-1
nest_coord[2,4]<-2

interaction$distance_nest<-NA
interaction$distance_food<-NA
interaction$nest_recipient<-factor(plyr::mapvalues(interaction$recipient,birdies,nests))

for(x in 1:nrow(interaction)){
  tmp<-c(interaction[x,c(13,14)])
  treat<-interaction[x,"treatment"]
  nest_coord[,treat]
  interaction[x,"distance_food"]<-dist(rbind(tmp,nest_coord[,treat]))
  
  treat<-interaction[x,"nest"]
  treat2<-interaction[x,"nest_recipient"]
  nest_coord[,treat]
  interaction[x,"distance_nest"]<-min(
    dist(rbind(tmp,nest_coord[,treat])),
    dist(rbind(tmp,nest_coord[,treat2])))
    
  
}

```

## Interaction plots
```{r interaction_counr}
interaction_count_tr<-interaction %>% dplyr::count(identity,treatment,exp_id, sort = TRUE)
interaction_count_cl<-interaction %>% dplyr::count(identity,closeness,exp_id, sort = TRUE)
#boxplot for interaction per treatment per indiviual
ggplot(interaction_count_tr, aes(x=treatment, y=n,colour=treatment))  +  #or distance_food
   geom_quasirandom()+theme_bw()+scale_color_viridis(discrete = T)+facet_wrap(. ~ identity)+theme(legend.position = 'top')+
  guides(color = guide_legend(title.position = 'top', title.hjust = .5, barwidth = unit(20, 'lines'), barheight = unit(.5, 'lines')))+ylab("Number of interactions") + labs(title="Interactions per treatment")


#plot of interaction counts normalized over exp_id per individual per treatment
interaction_count_cl$identity<-as.factor(interaction_count_cl$identity)
levels(interaction_count_cl$exp_id)<-sort(c(levels(interaction_count_cl$exp_id),"5-1"))

for (x in levels(interaction_count_cl$exp_id)){
  for(y in levels(interaction_count_cl$identity)){
    if(sum(interaction$identity==y & interaction$exp_id==x)==0){
      print(paste(x,y))
      tmp<-interaction_count_cl[interaction_count_cl$identity==y,][1,]
      tmp[,3]<-x
      tmp[,4]<-0
      interaction_count_cl<-rbind(interaction_count_cl,tmp)
    }
  }
}



#boxplot for interactions related to proximity of nest
ggplot(interaction_count_cl, aes(x=closeness, y=n,fill=closeness)) + 
  ## add half-violin from {ggdist} package
  geom_boxplot(
    width = .22, 
    ## remove outliers
    outlier.color = NA ## `outlier.shape = NA` works as well
  ) +
  ## add dot plots from {ggdist} package
  ggdist::stat_dots(
    ## orientation to the left
    side = "left", 
    ## move geom to the left
    justification = 1.3, 
    ## adjust grouping (binning) of observations 
    binwidth = .25
  )+
  theme_bw()+scale_fill_viridis(discrete = TRUE)+facet_wrap(. ~ identity)+theme(legend.position = 'top')+guides(color = guide_colorbar(title.position = 'top', title.hjust = .5,
 barwidth = unit(20, 'lines'), barheight = unit(.5, 'lines')))+
  coord_cartesian(xlim = c(1, NA), clip = "off") +
  labs(title="Interactions related to proximity of nest") +
ylab("number of interactions") + xlab("proximity to nest")
#dist(rbind(tmp,nest_coord[,1]))
#boxplot for interactions, distances per treatment
ggplot(interaction, aes(x=treatment, y=distance_nest,fill=treatment)) + #or distance_food
  geom_boxplot(outlier.colour="black", outlier.shape=16,
             outlier.size=2, notch=FALSE,show.legend=F) +
  theme_bw()+scale_fill_viridis(discrete = T)+theme(legend.position = 'top')+
  guides(fill = guide_legend(title.position = 'top',override.aes=list(size=10), title.hjust = .5,
 barwidth = unit(20, 'lines'), barheight = unit(.5, 'lines')))+
  coord_cartesian(xlim = c(1, NA), clip = "off") +
  labs(title="Interactions related to distance of the nest") +
xlab("treatment") + ylab("distance to nest")

#distance_food
ggplot(interaction, aes(x=treatment, y=distance_food,fill=treatment)) + #or distance_food
  geom_boxplot(outlier.colour="black", outlier.shape=16,
             outlier.size=2, notch=FALSE,show.legend=F) +
  theme_bw()+scale_fill_viridis(discrete = T)+theme(legend.position = 'top')+
  guides(fill = guide_legend(title.position = 'top',override.aes=list(size=10), title.hjust = .5,
 barwidth = unit(20, 'lines'), barheight = unit(.5, 'lines')))+
  coord_cartesian(xlim = c(1, NA), clip = "off") +
  labs(title="Interactions related to distance to food") +
xlab("treatment") + ylab("distance to food")
```
Creating heatmaps for interaction per treatment

```{r pressure, echo=FALSE}
interaction_count<-interaction %>% dplyr::count(x_pos,y_pos,treatment, sort = TRUE)
#Interaction plot per treatment
ggplot(interaction_count,aes(x=x_pos,y=y_pos,fill=n))+geom_tile( colour = "grey50")+scale_fill_viridis() + theme_bw()+facet_wrap(. ~ treatment)+
 labs(title="Interactions per treatment") +
ylab("y position") + xlab("x position")
#ggplotly(p)
```

```{r interaction plot,message=FALSE,warning=FALSE,fig.width=8, fig.height=9}
#Interaction plot per treatment per individual
interaction_count<-interaction %>% dplyr::count(x_pos,y_pos,treatment,identity, sort = TRUE)
ggplot(interaction_count,aes(x=x_pos,y=y_pos,fill=n))+geom_tile( colour = "grey50")+scale_fill_viridis() + theme_bw()+facet_grid(identity~treatment)+
 labs(title="Interactions per treatment per individual") +
ylab("y position") + xlab("x position")
```

```{r sex and session, echo=FALSE}
#including sex
interaction_count<-interaction %>% dplyr::count(x_pos,y_pos,treatment,sex, sort = TRUE)
ggplot(interaction_count,aes(x=x_pos,y=y_pos,fill=n))+geom_tile( colour = "grey50")+scale_fill_viridis() + theme_bw()+facet_grid(sex ~ treatment) +
 labs(title="Interactions per treatment related to sex") +
ylab("y position") + xlab("x position")
#ggplotly(p)
interaction$mod_ses<-0
for(x in levels(interaction$treatment)){
tmp<-interaction[interaction$treatment==x,]
tmp_factor<-factor(tmp$exp_id)
interaction[interaction$treatment==x,]$mod_ses<-as.numeric(tmp_factor)
}
interaction_count<-interaction %>% dplyr::count(x_pos,y_pos,treatment,mod_ses, sort = TRUE)
ggplot(interaction_count,aes(x=x_pos,y=y_pos,fill=n))+geom_tile( colour = "grey50")+scale_fill_viridis() + theme_bw()+facet_grid(treatment ~ mod_ses)+
 labs(title="Interactions per session") +
ylab("y position") + xlab("x position")


#ggplotly(p)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
```{r dist_test}
ks.test(interaction$distance_food,"punif",0,5)
```
